---
title: "Introduction to meta3l: Three-Level Meta-Analysis"
author: "meta3l package authors"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to meta3l: Three-Level Meta-Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

Three-level meta-analysis extends the classical two-level model to handle
**nested data**: multiple effect sizes reported within a single study. This
structure arises when studies measure the same outcome in several subsamples,
time points, or instruments. Ignoring the nesting inflates precision and
produces incorrect confidence intervals.

`meta3l` wraps the key steps of a three-level analysis into a single call:

- `metafor::escalc()` — compute effect sizes and sampling variances
- `metafor::vcalc()` — build a block-diagonal sampling covariance matrix
- `metafor::rma.mv()` — fit the multilevel random-effects model
- `clubSandwich::robust()` — apply cluster-robust standard errors

Supported effect size measures:

| Measure | Description |
|---------|-------------|
| `PLO`   | Proportion (logit-transformed) |
| `PAS`   | Proportion (arcsine-transformed) |
| `SMD`   | Standardized mean difference |
| `MD`    | Raw mean difference |
| `RR`    | Log risk ratio |
| `OR`    | Log odds ratio |

Back-transformation to the original scale is handled automatically.

## Installation

```{r eval=FALSE}
# From source (development version)
install.packages("meta3l", repos = NULL, type = "source")

# Or using remotes if hosted on GitHub
# remotes::install_github("user/meta3l")
```

## Quick Start: Fitting a Model

```{r eval=FALSE}
library(meta3l)

# Simulated dataset: prevalence of a condition across studies
dat <- data.frame(
  studlab = rep(c("Adams 2019", "Baker 2020", "Clark 2021", "Davis 2022"), each = 3),
  event   = c(12, 15, 10, 20, 18, 22, 8, 11, 9, 25, 19, 23),
  n       = c(100, 120, 95, 150, 130, 160, 80, 90, 85, 200, 170, 190),
  region  = rep(c("North", "South"), each = 6),
  year    = c(2019, 2019, 2019, 2020, 2020, 2020, 2021, 2021, 2021, 2022, 2022, 2022)
)

result <- meta3L(dat, slab = "studlab", event = "event", n = "n", measure = "PLO")
```

`meta3L()` treats each row as one effect size and each unique `studlab` value
as a cluster (study). The default cluster column is `studlab`; pass
`cluster = "author"` to use a different column.

## Examining Results

```{r eval=FALSE}
print(result)
summary(result)
```

Key components of the printed output:

- **Pooled estimate**: back-transformed (proportion for PLO/PAS, ratio for
  RR/OR, original units for SMD/MD) with 95% CI
- **k / n**: number of studies (clusters) and total number of effect sizes
- **I-squared decomposition**:
  - `I2_total`: total heterogeneity
  - `I2_between`: between-study variance as share of total variance
  - `I2_within`: within-study variance as share of total variance

Accessing individual components:

```{r eval=FALSE}
result$estimate    # back-transformed pooled estimate
result$ci.lb       # lower 95% confidence limit
result$ci.ub       # upper 95% confidence limit
result$i2          # named vector: total, between, within
result$model       # the underlying rma.mv object (full metafor access)
```

## Forest Plots

`forest.meta3L()` draws a publication-quality forest plot using grid graphics.

```{r eval=FALSE}
# Display in the viewer (no file saved)
forest.meta3L(result, file = NULL)

# Save with an auto-generated filename (meta3l_plot_YYYYMMDD_HHMMSS.png)
forest.meta3L(result)

# Add annotation columns for raw data
forest.meta3L(
  result,
  ilab     = c("event", "n"),
  ilab.lab = c("Events", "Total"),
  file     = NULL
)

# Save as PDF with an explicit path
forest.meta3L(result, file = "my_forest.pdf", format = "pdf")
```

Key arguments:

| Argument      | Description |
|---------------|-------------|
| `ilab`        | Character vector of column names to display to the left of the plot |
| `ilab.lab`    | Display labels for the `ilab` columns |
| `sortvar`     | Column name to sort rows by |
| `refline`     | Position of the reference line (default: null value for the measure) |
| `xlim`        | x-axis limits; auto-computed if `NULL` |
| `showweights` | Show study weights as a column |
| `file`        | Output file path (`NULL` = display only, `character(0)` = auto-name) |
| `format`      | `"png"` (default) or `"pdf"` |

## Subgroup Analysis

Use `moderator()` to test a categorical moderator. It fits a model with a
fixed effect for the subgroup variable and reports:

- Wald test for the moderator (from `robust()`)
- Likelihood ratio test comparing models with and without the moderator
- Per-subgroup pooled estimates on the back-transformed scale

```{r eval=FALSE}
mod_result <- moderator(result, subgroup = "region")
print(mod_result)
```

To visualise subgroup-specific pooled estimates alongside individual studies:

```{r eval=FALSE}
forest_subgroup.meta3L(result, subgroup = "region", file = NULL)
```

The subgroup forest plot groups rows by level of `subgroup`, inserts a
diamond for each group's pooled estimate, and adds an overall diamond at
the bottom.

## Meta-Regression: Continuous Moderator

`bubble.meta3L()` fits a meta-regression with a continuous moderator and
draws a bubble plot where bubble sizes are proportional to study precision.

```{r eval=FALSE}
bub <- bubble.meta3L(result, mod = "year", file = NULL)

# Summary of the meta-regression model
bub$summary
# Returns: estimate (back-transformed at mean moderator), CI, R-squared,
#          slope, robust p-value from clubSandwich
```

The plot shows the back-transformed regression line with a 95% confidence
band over the observed range of the moderator.

## Sensitivity Analysis: Leave-One-Out

`meta3l` provides two leave-one-out (LOO) functions that exploit the three-
level structure:

**Cluster-level LOO** — drop one study (cluster) at a time:

```{r eval=FALSE}
loo_c <- loo_cluster(result, file = NULL)

# Table of results: one row per dropped study
loo_c$table
# Columns: dropped study, k remaining, pooled estimate, CI, I2_total
```

**Effect-level LOO** — drop one effect size at a time:

```{r eval=FALSE}
loo_e <- loo_effect(result, file = NULL)

# Table of results: one row per dropped effect size
loo_e$table
```

Both functions produce an influence plot (estimate with CI for each leave-
one-out iteration, reference line at the full-model estimate). Large shifts
in the pooled estimate or I-squared when a single study is dropped indicate
that the result is sensitive to that observation and warrants closer
inspection.

Pass `file = NULL` to display the plot without saving; omit `file` or set
`file = character(0)` for automatic file naming.

## Working with Excel Data

`read_multisheet_excel()` reads every sheet from a multi-sheet Excel workbook
and returns a named list of data frames.

```{r eval=FALSE}
sheets <- read_multisheet_excel("path/to/data.xlsx")

# Returns a named list — one element per sheet
names(sheets)          # sheet names
sheets[["Sheet1"]]     # access individual sheet as a data frame
```

This is useful when effect sizes from different subgroups or outcomes are
stored on separate worksheets that need to be combined before analysis.

## Summary

The typical meta3l workflow is:

1. **Fit**: `result <- meta3L(dat, slab, event, n, measure)`
2. **Inspect**: `print(result)` / `summary(result)`
3. **Visualise**: `forest.meta3L(result)`
4. **Subgroup**: `moderator(result, subgroup)` + `forest_subgroup.meta3L(result, subgroup)`
5. **Regression**: `bubble.meta3L(result, mod)`
6. **Sensitivity**: `loo_cluster(result)` / `loo_effect(result)`

All six effect size measures (`PLO`, `PAS`, `SMD`, `MD`, `RR`, `OR`) follow
the same workflow — meta3l selects the correct escalc call and applies the
appropriate back-transformation automatically.

For full argument documentation see `?meta3L`, `?forest.meta3L`,
`?moderator`, `?bubble.meta3L`, `?loo_cluster`, `?loo_effect`, and
`?read_multisheet_excel`.
